---
title: 'Today project'
author: "Hongxu Wang"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---
<style type="text/css">
  body{
  font-size: 12pt;
}
</style>

```{r exposure outcome setup, include=FALSE}
source(here::here("!libraries.R"))
source(here::here("!directories.R"))
source(here::here("!load_clean_data.R"))
source(here::here("!functions.R"))
source(here::here("analysis function_Multiple exposures and multiple outcomes analysis.R"))

options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r define variables}
outcome_glu <- c("codi", "si_1_ins0", "hb_a1c")

outcome_biomaker <- c("seq.2836.68",
                      "seq.9021.1",
                      "seq.5661.15",
                      "seq.11516.7",
                      "seq.17138.8",
                      "seq.12446.49",
                      "seq.15509.2")
pfas_min <-  c("pfas_pfna", "pfas_pfda", "pfas_pfhpa", 'pfas_pfoa', 
               "pfas_pfos", "pfas_pfhps", "pfas_pfhxs")

pfas_names_all <- c("pfas_pfna", 
                    "pfas_pfda", 
                    "pfas_pfhpa", 
                    "pfas_pfoa",
                    "pfas_pfos", 
                    "pfas_pfhps",
                    "pfas_pfhxs",
                    "pfas_nmefosaa", 
                    "pfas_pfuna"
                    )

covars <- c("case1_control0", 
            "sex", 
            "agebase", 
            # kidney functions
            # "u_alb_creat",
            # "uacid",
            # "est_creat_clear",
            "serum_creat"
            )

levels <- c("PFUnA", "PFHpS","PFHpA","NMeFOSAA","PFDA","PFNA","PFHxS", "PFOA", "PFOS"
)
```

## Descriptive PFAS
```{r descriptive pfas}
# long format data for plasma pfas
data_l <- full_data %>% 
  pivot_longer(
  names_to = "pfas",
  values_to = "value",
  cols = c(all_of(pfas_names_all)))

# descriptive statistics
stats <- data_l %>%
  drop_na(value) %>%
  group_by(pfas) %>%
  dplyr::summarise(
    geometric_mean = fungm(value),
    percentile_50 = qntle_fxn(value, .50),
    percentile_75 = qntle_fxn(value, .75),
    percentile_90 = qntle_fxn(value, .9)) %>%
  ungroup()

# Rename columns
stats_summary <- stats %>%
  dplyr::rename(
    "PFAS" = pfas,
    "Geometric Mean [95% CI]" = geometric_mean,
    "50th percentile" = percentile_50,
    "75th percentile" = percentile_75,
    "90th percentile" = percentile_90)
# save result
write_csv(stats_summary, fs::path(dir_results,
                                  "1_0_descriptive_stats_pfas.csv"))
stats_summary
```

## Correlation of the PFAS
```{r correlation of pfas}
cor <- cor(full_data %>% dplyr::select(pfas_names_all))
corrplot::corrplot(cor)
```

## Exposure burden score
```{r exposure burden score}
data_analysis <- full_data %>%
  mutate_at(.vars = vars(all_of(pfas_names_all[c(1,4,5,8)])), 
            .funs = list(
                         decile = ~as.integer(cut(., quantile(., probs = seq(0, 1, 0.1)), include.lowest = TRUE
            )))) %>%
  mutate(pfas_pfhxs_decile = cut(pfas_pfhxs, 
                                 quantile(pfas_pfhxs, probs = seq(0, 1, 1/9)), include.lowest = TRUE) %>% as.integer()) %>%
  mutate(pfas_pfda_detected=
           ifelse(pfas_pfda<0.05|pfas_pfda==0.2/sqrt(2), 1, 2)) %>%
  # mutate(pfas_pfhps_detected =
  #          ifelse(pfas_pfhps<0.1|pfas_pfhps==0.2/sqrt(2), 1, 2)) %>%
   mutate_at(.vars = vars(c("pfas_pfhpa","pfas_pfuna")),
            .funs = list(detected = ~ifelse(.<0.05|.==0.1/sqrt(2), 1, 2))) %>%
  mutate_at(.vars = vars(c("pfas_pfhpa", "pfas_pfhps")), 
            .funs = list(tertile = ~cut(., 
                                 quantile(., probs = seq(0, 1, 1/3)), include.lowest = TRUE) %>% as.integer()))


burden_score_pfas<- data_analysis %>% 
                                  dplyr::select(contains("detected"), contains("decile"), contains("tertile"), -pfas_pfhpa_detected) %>% as.data.frame()

eap.d <- ltm::factor.scores(grm(burden_score_pfas), 
                            method="EAP", resp.patterns =
                              burden_score_pfas)$score.dat$z1

plot(grm(burden_score_pfas), type = "IIC",legend = TRUE, xlab = "PFAS Burden", main = "", cx = "topright", cex = 0.6) 

data_analysis1 <- data_analysis %>% bind_cols(score = eap.d)

data <- data_analysis1 %>% 
  fastDummies::dummy_cols(select_columns = c('sex'),
                          remove_selected_columns = FALSE, 
                          remove_most_frequent_dummy = TRUE) |>
  mutate(
         yrstoglyc = (daystoglyc)/365,
         yrstoglyc2 = yrstoglyc + dxtime/12,
         yrstohyp = (daystohyp)/365, 
         yrstohyp2 = yrstohyp + dxtime/12, 
         yrstomic = (daystomic)/365, 
         yrstomic2 = (yrstomic+dxtime/12),
         yrstomac = (daystomac)/365,
         yrstomac2 = (yrstomac+dxtime/12)
         # mac2 = if_else(yrstomac > 8.5, 0, 1),
         # mic2 = if_else(yrstomic > 8.5, 0, 1), 
         # hyp2 = if_else(yrstohyp > 8.5, 0, 1)
         ) %>%
  mutate_at(.vars = vars(c(all_of(pfas_names_all), score)),
            .funs = list(median = ~ifelse(.<median(.),"0", "1")
                         )) %>%
  mutate(score_tertile = cut(score, 
                             quantile(score, probs = seq(0, 1, 1/3)),
                             include.lowest = TRUE) %>% 
           as.integer() %>% as.character(),
         score_quartile = cut(score,
                             quantile(score, probs = seq(0, 1, 1/4)),
                              include.lowest = TRUE) %>%
           as.integer() %>% as.character(),
         score_quintile = cut(score,
                             quantile(score, probs = seq(0, 1, 1/5)),
                              include.lowest = TRUE) %>%
           as.integer() %>% as.character()) %>%
  mutate_at(.vars = vars(all_of(pfas_names_all)),
            .funs = ~log2(.) %>% scale(.) %>% as.numeric(.))%>%
  mutate_at(.vars = vars(c(outcome_glu, outcome_biomaker)),
            .funs = ~scale(.) %>% as.numeric(.))
```


## Analysis with whole group 
**adjusted for "case1_control0", "sex", "agebase", "serum_creat"**
```{r whole group}
  
pfas_names_all1 <- c(pfas_names_all, "score")

cont_outcome_result <- model_output(pfas_names_all1,
                                    outcomes = c(outcome_glu, outcome_biomaker),
                                    covars = covars,
                                    outcome_family = "gaussian",
                                    data = data) %>%
  mutate(sig = ifelse(p.value < 0.05, "Sig.", "Not Sig."))

# cont_outcome_result <- epiomics::owas(df = data,
#                                       var = c(outcome_glu, outcome_biomaker),
#                                       omics = pfas_names_all,
#                                       covars = covars,
#                                       var_exposure_or_outcome = "outcome",
#                                       conf_int = TRUE)

cont_result <- cont_outcome_result %>% 
  filter(term == "exposure_concentration") %>%
  dplyr::select(-term) %>%
  rename(pfas = exposures) %>%
  rename_pfas()
  
## coefficient plot
(p1 <- cont_result %>%
  filter(outcome %in% c(outcome_glu)) %>%
  ggplot(aes(x = factor(pfas, levels = c(levels,"score")),y = estimate, color = sig)) +
  geom_point(size = 1) +
  geom_errorbar(aes(ymin = conf.low,
                    ymax = conf.high),
                width = 0) +
  geom_hline(yintercept = 0, linetype = 2) +
  facet_grid( ~ outcome, scales = "free") +
  ylab("Coefficient (95% CI)") +
  # ggtitle("Coefficient Plot (Binary Logistic Regression)") +
  theme(text = element_text(size = 10),
        axis.title.y = element_blank(),
        panel.background = element_rect(fill="white"), 
        strip.background = element_rect(fill = "white"),
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        legend.position = "none") +
  coord_flip() +
  scale_color_manual(values = c("grey", "red")))

ggsave(fs::path(dir_figure, 
                "1127_hw",
                "coef_plot_glucose_outcome.png"), bg = "white",
       width = 8, height = 3, dpi = 300)

(p2 <- cont_result %>%
  filter(outcome %in% c(outcome_biomaker)) %>%
  ggplot(aes(x = factor(pfas, levels = c(levels,"score")),y = estimate, color = sig)) +
  geom_point(size = 1) +
  geom_errorbar(aes(ymin = conf.low,
                    ymax = conf.high),
                width = 0) +
  geom_hline(yintercept = 0, linetype = 2) +
  facet_grid( ~ outcome, scales = "free") +
  ylab("Coefficient (95% CI)") +
  # ggtitle("Coefficient Plot (Binary Logistic Regression)") +
  theme(text = element_text(size = 10),
        axis.title.y = element_blank(),
        panel.background = element_rect(fill="white"), 
        strip.background = element_rect(fill = "white"),
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        legend.position = "none") +
  coord_flip() +
  scale_color_manual(values = c("grey", "red")))

ggsave(fs::path(dir_figure, 
                "1127_hw",
                "coef_plot_biomarker_outcome.png"), bg = "white",
       width = 8, height = 3, dpi = 300)
```

## Survival Analysis -- Individual PFAS

```{r survival individual pfas}
# try different kidney function: 

ind_vars <- tibble(exposure = colnames(data %>%
                                         dplyr::select(contains("median"), contains("score"))))

dep_vars <- tibble(time = c("yrstoglyc2","yrstohyp2",
                            "yrstomic2"), 
                   event = c("glyc", "hyp", "mic"))

eo_comb <- list(pfas = ind_vars$exposure, event = dep_vars$event) %>%
  cross_df() %>% left_join(dep_vars, by = "event")

models <- eo_comb %>%
  mutate(covar = paste0("sex_male+",str_c(covars[-2], collapse = "+")),
         formula = str_c("Surv(", time , ",", event, ")", "~", pfas, "+", covar))

models$output <- map(models$formula,
                     ~coxph(as.formula(.),
                            data = data) %>%
                       tidy(., conf.int = TRUE))

mod_output_df <- models %>%
  unnest(output) %>%
  filter(grepl("median", term)|grepl("score", term)&!grepl("quintile",term))%>%
  mutate(HR = exp(estimate),
         exp_conf_high = exp(conf.high),
         exp_conf_low = exp(conf.low),
         sig = ifelse(p.value < 0.05, "Sig.", "Not Sig.")) %>%
  rename_pfas()%>%
  mutate(pfas = ifelse(pfas %in% levels, pfas, term))
# Note: 1 unit increase in pfas resulting in xx-fold increase in hazard.

(p <- mod_output_df %>%
  # filter(grepl("score", pfas)) %>%
  ggplot(aes(x = pfas,y = estimate, color = sig)) +
  geom_point(size = 1) +
  geom_errorbar(aes(ymin = conf.low,
                    ymax = conf.high),
                width = 0) +
  geom_hline(yintercept = 0, linetype = 2) +
  facet_grid( ~ event, scales = "free") +
  ylab("Log HR (95% CI)") +
  # ggtitle("Coefficient Plot (Binary Logistic Regression)") +
  theme(text = element_text(size = 10),
        axis.title.y = element_blank(),
        panel.background = element_rect(fill="white"), 
        strip.background = element_rect(fill = "white"),
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        legend.position = "none") +
  coord_flip() +
  scale_color_manual(values = c("grey", "red")))

ggsave(fs::path(dir_figure, 
                "1127_hw",
                "Hazard_ratio_burden_score_serum_creat.png"), bg = "white",
       width = 8, height = 3, dpi = 300)
```

## Survival Analysis - Burden score(Binary) and hyperfiltration

```{r survival analysis}
df = data %>% as.data.frame()

fit <- coxph(Surv(yrstohyp2, hyp) ~ 
                           strata(score_median) + case1_control0 + sex + agebase + serum_creat, data = df)

# adjustedsurv(data = df,
#              variable = "score_median",
#              ev_time = "yrstohyp2",
#              event = "hyp",
#              method = "direct",
#              outcome_model = fit,
#              conf_int = TRUE, )
ggadjustedcurves(fit,
                 data = df,
                 method = "conditional",
                 variable = "score_median",
                 # conf.int = TRUE,
                 # pval = TRUE,
                 # risk.table = TRUE,
                 # legend.labs = c("Below Median", "Above Median"),
                 legend.title = "PFAS Burden Score",
                 palette = c("dodgerblue2", "orchid2"),
                 title = "Kaplan-Meier Curve",
                 # risk.table.height = 0.3
                 )

ggsave(filename = fs::path(dir_figure,"1127_hw", "KM_curve_score_hyp.jpeg"), 
       width = 5, 
       height = 5, dpi = 3000)
```


## Mixture analysis - qgcomp
```{r mixture analysis}
# qgcomp --------
qgcomp <- owas_qgcomp(df = data, 
                           expnms = pfas_names_all, 
                           omics = c(outcome_glu, outcome_biomaker), 
                           covars = covars,
                           confidence_level = 0.95,
                           q = 3) |>
  dplyr::select(feature, psi, lcl_psi, ucl_psi, p_value) %>%
  mutate(type = ifelse(feature %in% outcome_glu, "clinical", "biomarker")) %>%
  mutate(feature = rename_outcomes(feature))


reorder_outcome_name <- c("Hyperglycemia (HbA1c)", 
                          "β-cell function (C-Peptide oDI)",  
                          "Ins. Sensitivity (1/fast. ins)",
                          "NGAL","KIM-1", "IL18",
                          "FABPL","a-GST", "a-GST1","NAG"
                          ) 
# Get color scale
colors <- RColorBrewer::brewer.pal(n=3, name = "Dark2")[1:2]

# Reorder outcome name and set color scale
qgcomp_res <- qgcomp %>%
  mutate(feature = factor(feature,
                               levels = reorder_outcome_name) %>%
           fct_rev(),
         color = if_else(type == "biomarker", 
                         colors[1], colors[2])) |>
  arrange(feature)

# Plot all outcomes 
(plotout <- qgcomp_res |>
    ggplot(aes(y = psi, 
               x = feature,
               color = color)) +
    geom_point(size = 3) +
    geom_errorbar(aes(ymin = lcl_psi,
                      ymax = ucl_psi),
                  width = 0, size = .75) +
    geom_hline(yintercept = 0, linetype = 2, linewidth = .5) +
    theme(axis.title.y = element_blank(), 
          axis.text.y = element_text(size = 14, vjust = .5)) +
    scale_color_manual(values = colors) +
    ylab("PFAS Mixture (95% CI)") + 
    coord_flip() + 
    theme(legend.position = "None",
          axis.text.y = element_text(color = qgcomp_res$color, 
                                     face = "bold")))

ggsave(plotout,
       filename = fs::path(dir_figure, 
                           "QGComp pfas mixture TODAY_1204.jpeg"),
       height = 2.25, width = 5, bg = "white")
```


## Survival analysis- Mixture PFAS
### glycemic failure
```{r glycemic failure}
(qc.survfit <- qgcomp::qgcomp.cox.noboot(
  survival::Surv(yrstoglyc2, glyc) ~ ., 
  expnms = pfas_names_all,
  data = data |> 
    dplyr::select(all_of(pfas_names_all), 
           covars[-2],
           sex_male,
           yrstoglyc2, 
           glyc),
  q = 2))
```
### hyperfiltration
```{r hyp}
(qc.survfit <- qgcomp::qgcomp.cox.noboot(
  survival::Surv(yrstohyp2, hyp) ~ ., 
  expnms = pfas_names_all,
  data = data |> 
    dplyr::select(all_of(pfas_names_all), 
           covars[-2],
           sex_male,
           yrstohyp2, 
           hyp),
  q = 2))
```

### microalbuminuria
```{r microalbuminuria}
(qc.survfit <- qgcomp::qgcomp.cox.noboot(
  survival::Surv(yrstomic2, mic) ~ ., 
  expnms = pfas_names_all,
  data = data |> 
    dplyr::select(all_of(pfas_names_all),
           covars[-2],
           sex_male,
           yrstomic2, 
           mic),
  q = 2))

```

## Causal Mediation Analysis with Survival Outcome
* Exposure: PFAS burden score
* Mediator: HbA1c
* Survival Outcome: Hyperfiltration (Kidney injury)
```{r CMA}
data1 <- data %>% mutate(score_median = factor(score_median))

mediation_result <- CMAverse::cmest(data1, 
                          outcome = "yrstohyp2", 
                          event = "hyp",
                          exposure = "score_median", 
                          mediator = c("hb_a1c"), 
                          basec = covars,
                          EMint = TRUE, 
                          mreg = list("linear"), 
                          yreg = "coxph", 
                          model = "gformula",
                          astar = 0, 
                          a = 1, 
                          mval = list(0), 
                          estimation = "imputation",
                          inference = "bootstrap",
                          nboot = 10)

summary(mediation_result)

CMAverse::ggcmest(mediation_result) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))
```

## High dimentionale mediation analysis - HIMA
* Exposure: PFAS burden score
* Mediator: All metabolites and protein biomakers
* Survival Outcome: Hyperfiltration (Kidney injury)

```{r hima}
met_names <- colnames(data %>% dplyr::select(contains("met"), contains("seq")))

var_uni_value <- data %>% 
  pivot_longer(cols = met_names) %>%
  group_by(name) %>%
  summarise(n = length(unique(value))) %>%
  filter(n == 1)

X1 = data$score
Z1 = data[, c(covars[-2], "sex_male")] %>% as.matrix()

M = data %>% dplyr::select(contains("met"),contains("seq"),
                    -var_uni_value$name) %>%as.matrix()
M1 = data %>% dplyr::select(contains("met"), -var_uni_value$name) %>%as.matrix()
M2 = data %>% dplyr::select(contains("seq")) %>% as.matrix()

(result_both<- HIMA::survHIMA(X = X1,
                        Z = Z1,
                        M = M,
                        OT = data$yrstohyp2,
                        status = data$hyp,
                        FDRcut = 1,
                        scale = TRUE,
                        verbose = FALSE))

(result_met<- HIMA::survHIMA(X = X1,
                        Z = Z1,
                        M = M1,
                        OT = data$yrstohyp2,
                        status = data$hyp,
                        FDRcut = 1,
                        scale = TRUE,
                        verbose = FALSE))

(result_pro<- HIMA::survHIMA(X = X1,
                        Z = Z1,
                        M = M2,
                        OT = data$yrstohyp2,
                        status = data$hyp,
                        FDRcut = 1,
                        scale = TRUE,
                        verbose = FALSE))


intersect(result_both$ID, outcome_biomaker)  
```

