---
title: 'Today project'
author: "Hongxu Wang"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---
<style type="text/css">
  body{
  font-size: 12pt;
}
</style>

```{r exposure outcome setup, include=FALSE}
source(here::here("!libraries.R"))
source(here::here("!directories.R"))
source(here::here("!load_clean_data.R"))
source(here::here("analysis function_Multiple exposures and multiple outcomes analysis.R"))

options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Linear regression stratified by cases vs. controls
*PFAS: log2 transformed
*Outcome: Scaled
```{r linear regression stratified by cases vs. controls}
outcome <- c("time_to_glyc_scld", 
             "codi", 
             "si_1_ins0",
             "hb_a1c", 
             "mic", 
             "daystomic", 
             # "mac",
             "daystomac", 
             "rapid",
             "daystorapid", 
             "hyp0", 
             "daystohyp")
## MAC only has 0 in that, so I exclude it

full_data <- full_data %>% 
  mutate_at(.vars = vars(outcome[c(1:4, 6,7,9, 11)]),
            .funs = ~scale(.))

# Other outcomes_HW -----
# this is a function to run owas
owas_mul <- function(outcome, family){
  # filter the full data to only include cases
  exwas_lm_case <- full_data |>
    tidylog::filter(case1_control0 == 1) |>
    # log2 transform the pfas variables
    mutate(across(all_of(pfas_names), log2)) |>
    mutate(across(all_of(pfas_names), scale)) |>
    # run the owas function from the epiomics package
    epiomics::owas(var = outcome,
                   var_exposure_or_outcome = "outcome",
                   covars = c("sex", "agebase", "tx"),
                   omics = pfas_names, 
                   family = family,
                   conf_int = TRUE)
  # if the family is gaussian, then calculate the confidence intervals for the beta estimates
  if(family == "gaussian"){
    exwas_lm_case <- exwas_lm_case %>% 
      mutate(beta_ci = jag2::effest_ci(estimate, conf_low, conf_high)) %>%
      mutate(case_type = "case") %>%
      select(case_type, everything())
    # if the family is binomial, then calculate the confidence intervals for the odds ratios
  }else{
    exwas_lm_case <- exwas_lm_case %>%
      mutate(odds_ratio_ci = jag2::effest_ci(exp(estimate), exp(conf_low), exp(conf_high)))%>%
      mutate(case_type = "case") %>%
      select(case_type, everything())
  }
  # filter out the pfas variables that are not included in the analysis
  exwas_lm_case <- exwas_lm_case %>% 
    filter(!feature_name %in% 
             c("pfas_br_pfos", "pfas_l_pfoa", "pfas_l_pfos")) %>%
    # create a new variable that is the name of the pfas variable
    mutate(pfas_name = case_when(grepl("pfos", feature_name) ~ "PFOS",
                                 grepl("pfoa", feature_name) ~ "PFOA",
                                 grepl("pfda", feature_name) ~ "PFDA",
                                 grepl("pfhpa", feature_name) ~ "PFHpA",
                                 grepl("pfhps", feature_name) ~ "PFHpS",
                                 grepl("pfhxs", feature_name) ~ "PFHxS",
                                 grepl("pfna", feature_name) ~ "PFNA",
                                 grepl("pfuna", feature_name) ~ "PFUnA",
                                 grepl("nmefosaa", feature_name) ~ "NMeFOSAA"))
  
  # filter the full data to only include controls
  exwas_lm_control <- full_data |>
    tidylog::filter(case1_control0 == 0) |>
    # log2 transform the pfas variables
    mutate(across(all_of(pfas_names), log2)) |>
    mutate(across(all_of(pfas_names), scale)) |>
    # run the owas function from the epiomics package
    epiomics::owas(var = outcome,
                   var_exposure_or_outcome = "outcome",
                   covars = c("sex", "agebase", "tx"),
                   omics = pfas_names, 
                   family = family,
                   conf_int = TRUE) 
  # if the family is gaussian, then calculate the confidence intervals for the beta estimates
  if(family == "gaussian"){
    exwas_lm_control <- exwas_lm_control %>% 
      mutate(beta_ci = jag2::effest_ci(estimate, conf_low, conf_high)) %>%
      mutate(case_type = "control") %>%
      select(case_type, everything())
    # if the family is binomial, then calculate the confidence intervals for the odds ratios
  }else{
    exwas_lm_control <- exwas_lm_control %>%
      mutate(odds_ratio_ci = jag2::effest_ci(exp(estimate), exp(conf_low), exp(conf_high)))%>%
      mutate(case_type = "control") %>%
      select(case_type, everything())
  }
  
  # filter out the pfas variables that are not included in the analysis
  exwas_lm_control <- exwas_lm_control %>% 
    filter(!feature_name %in% 
             c("pfas_br_pfos", "pfas_l_pfoa", "pfas_l_pfos")) %>%
    # create a new variable that is the name of the pfas variable
    mutate(pfas_name = case_when(grepl("pfos", feature_name) ~ "PFOS",
                                 grepl("pfoa", feature_name) ~ "PFOA",
                                 grepl("pfda", feature_name) ~ "PFDA",
                                 grepl("pfhpa", feature_name) ~ "PFHpA",
                                 grepl("pfhps", feature_name) ~ "PFHpS",
                                 grepl("pfhxs", feature_name) ~ "PFHxS",
                                 grepl("pfna", feature_name) ~ "PFNA",
                                 grepl("pfuna", feature_name) ~ "PFUnA",
                                 grepl("nmefosaa", feature_name) ~ "NMeFOSAA"))
  
  # bind the case and control data together
  result <- exwas_lm_case %>% bind_rows(exwas_lm_control)
  return(result)
}


# Categorical outcomes
mic_result    <- owas_mul("mic",   "binomial")
rapid_result  <- owas_mul("rapid", "binomial")
hyp0_result   <- owas_mul("hyp0",  "binomial")
result_cat <- mic_result  %>% bind_rows(rapid_result) %>%
  bind_rows(hyp0_result) 


# Continuous outcomes
time_to_glyc_scld_result    <- owas_mul("time_to_glyc_scld", "gaussian")
codi_result                 <- owas_mul("codi",              "gaussian")
si_1_ins0_result            <- owas_mul("si_1_ins0",         "gaussian")
hb_a1c_result               <- owas_mul("hb_a1c",            "gaussian")
daystomic_result   <- owas_mul("daystomic",   "gaussian")
daystomac_result   <- owas_mul("daystomac",   "gaussian")
daystorapid_result <- owas_mul("daystorapid", "gaussian")
daystohyp_result   <- owas_mul("daystohyp",   "gaussian")

result_cont<- daystomic_result %>%
  bind_rows(time_to_glyc_scld_result)  |>
  bind_rows(codi_result)  |>
  bind_rows(si_1_ins0_result)  |>
  bind_rows(hb_a1c_result)  |>
  bind_rows(daystomac_result) %>% 
  bind_rows(daystorapid_result) %>% 
  bind_rows(daystohyp_result)

# Combine continuous and categorical
result <- result_cat %>% bind_rows(result_cont)

write_csv(result,
          fs::path(dir_results,
                   "Stratified case control analysis multiple outcomes 1127.csv"))

levels <- c("PFUnA", "PFHpS","PFHpA","NMeFOSAA","PFDA","PFNA","PFHxS", "PFOA", "PFOS" 
)
```

### Figure1. Coefficient plot
**The estimates of categorical outcome are too large. Categorical outcomes were excluded from following analysis **
```{r plot}
## coefficient plot for rapid progression
(p1 <- result_cont %>%
  ggplot(aes(x = factor(pfas_name, levels = levels),y = estimate, color = threshold)) +
  geom_point(size = 1) +
  geom_errorbar(aes(ymin = conf_low,
                    ymax = conf_high),
                width = 0) +
  geom_hline(yintercept = 0, linetype = 2) +
  facet_grid( case_type ~ var_name , scales = "free") +
  ylab("Coefficient (95% CI)") +
  # ggtitle("Coefficient Plot (Binary Logistic Regression)") +
  theme(text = element_text(size = 10),
        axis.title.y = element_blank(),
        panel.background = element_rect(fill="white"), 
        strip.background = element_rect(fill = "white"),
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        legend.position = "none") +
  coord_flip() +
  scale_color_manual(values = c("grey", "red")))

# ggsave(fs::path(dir_figure,
#                    "1127_hw",
#                    "coef_plot_cont_outcome_1127.png"),
#        bg = "white",
#        width = 12, height = 3, dpi = 300)

(p2 <- result_cat %>%
  ggplot(aes(x = factor(pfas_name, levels = levels),y = estimate, color = threshold)) +
  geom_point(size = 1) +
  coord_flip() +
  geom_errorbar(aes(ymin = conf_low,
                    ymax = conf_high),
                width = 0) +
  geom_hline(yintercept = 1, linetype = 2) +
  facet_grid( case_type ~ var_name , scales = "free") +
  ylab("Log Odds Ratio (95% CI)") +
  theme(text = element_text(size = 10),
        axis.title.y = element_blank(),
        panel.background = element_rect(fill="white"), 
        strip.background = element_rect(fill = "white"),
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        legend.position = "none") +
  coord_flip() +
  scale_color_manual(values = c("grey", "red"))) 
```

## Analysis with whole group adjusted for "case1_control0", "sex", "agebase", "tx"
```{r whole group}
# Analysis with whole group adjusted for "case1_control0", "sex", "agebase", "tx"--------
data <- full_data %>% 
  mutate(across(all_of(pfas_names), log2)) |>
  mutate(across(all_of(pfas_names), scale)) |>
  mutate_at(.vars = vars(outcome[c(1:4, 6,7,9, 11)]),
            .funs = ~scale(.))

cont_outcome_result <- model_output(pfas_names,
                                    outcomes = outcome[c(1:4, 6,7,9,11)],
                                    covars = c("case1_control0", "sex", "agebase", "tx") ,
                                    outcome_family = "gaussian",
                                    data = data) %>%
  mutate(sig = ifelse(p.value < 0.05, "Sig.", "Not Sig."))

cont_result <- cont_outcome_result %>% 
  filter(term == "exposure_concentration") %>%
  select(-term) %>%
  filter(!exposures %in% 
           c("pfas_br_pfos", "pfas_l_pfoa", "pfas_l_pfos")) %>%
  mutate(pfas_name = case_when(grepl("pfos", exposures) ~ "PFOS",
                               grepl("pfoa", exposures) ~ "PFOA",
                               grepl("pfda", exposures) ~ "PFDA",
                               grepl("pfhpa", exposures) ~ "PFHpA",
                               grepl("pfhps", exposures) ~ "PFHpS",
                               grepl("pfhxs", exposures) ~ "PFHxS",
                               grepl("pfna", exposures) ~ "PFNA",
                               grepl("pfuna", exposures) ~ "PFUnA",
                               grepl("nmefosaa", exposures) ~ "NMeFOSAA"))

cat_outcome_result <- model_output(pfas_names,
                                   outcomes = outcome[c(5, 8, 10)],
                                   covars = c("case1_control0", "sex", "agebase", "tx"),
                                   outcome_family = "binomial",
                                   data = data)

cat_result <- cat_outcome_result %>% 
  filter(term == "exposure_concentration") %>%
  select(-term) %>%
  filter(!exposures %in% 
           c("pfas_br_pfos", "pfas_l_pfoa", "pfas_l_pfos")) %>%
  mutate(pfas_name = case_when(grepl("pfos", exposures) ~ "PFOS",
                               grepl("pfoa", exposures) ~ "PFOA",
                               grepl("pfda", exposures) ~ "PFDA",
                               grepl("pfhpa", exposures) ~ "PFHpA",
                               grepl("pfhps", exposures) ~ "PFHpS",
                               grepl("pfhxs", exposures) ~ "PFHxS",
                               grepl("pfna", exposures) ~ "PFNA",
                               grepl("pfuna", exposures) ~ "PFUnA",
                               grepl("nmefosaa", exposures) ~ "NMeFOSAA")) %>%
  mutate(sig = ifelse(p.value < 0.05, "Sig.", "Not Sig."))


## coefficient plot
(p1 <- cont_result %>%
  ggplot(aes(x = factor(pfas_name, levels = levels),y = estimate, color = sig)) +
  geom_point(size = 1) +
  geom_errorbar(aes(ymin = conf.low,
                    ymax = conf.high),
                width = 0) +
  geom_hline(yintercept = 0, linetype = 2) +
  facet_grid( ~ outcome, scales = "free") +
  ylab("Coefficient (95% CI)") +
  # ggtitle("Coefficient Plot (Binary Logistic Regression)") +
  theme(text = element_text(size = 10),
        axis.title.y = element_blank(),
        panel.background = element_rect(fill="white"), 
        strip.background = element_rect(fill = "white"),
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        legend.position = "none") +
  coord_flip() +
  scale_color_manual(values = c("grey", "red")))

ggsave(fs::path(dir_figure, 
                "1127_hw",
                "coef_plot_cont_outcome_full_group.png"), bg = "white",
       width = 8, height = 3, dpi = 300)

(p2 <- cat_result %>%
  ggplot(aes(x = factor(pfas_name, levels = levels),y = exp(estimate), color = sig)) +
  geom_point(size = 1) +
  coord_flip() +
  geom_errorbar(aes(ymin = exp(conf.low),
                    ymax = exp(conf.high)),
                width = 0) +
  geom_hline(yintercept = 1, linetype = 2) +
  facet_grid( ~ outcome, scales = "free") +
  ylab("Odds Ratio (95% CI)") +
  theme(text = element_text(size = 10),
        axis.title.y = element_blank(),
        panel.background = element_rect(fill="white"), 
        strip.background = element_rect(fill = "white"),
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        legend.position = "none") +
  coord_flip() +
  scale_color_manual(values = c("grey", "red")))

ggsave(fs::path(dir_figure, 
                "1127_hw",
                "coef_plot_cat_outcome_full_group.png"), bg = "white",
       width =  8, height = 3, dpi = 300)
```

## Analysis with whole group adjusted for "case1_control0", "sex", "agebase", "tx", "est_creat_clear"
```{r extra covars}
cont_outcome_result1 <- model_output(pfas_names,
                                     outcomes = outcome[c(1:4, 6,7,9,11)],
                                     covars = c("case1_control0", "sex", "agebase", "tx", "est_creat_clear") ,
                                     outcome_family = "gaussian",
                                     data = data) 
cont_result1 <- cont_outcome_result1 %>% 
  filter(term == "exposure_concentration") %>%
  select(-term)%>%
  filter(!exposures %in% 
           c("pfas_br_pfos", "pfas_l_pfoa", "pfas_l_pfos")) %>%
  mutate(pfas_name = case_when(grepl("pfos", exposures) ~ "PFOS",
                               grepl("pfoa", exposures) ~ "PFOA",
                               grepl("pfda", exposures) ~ "PFDA",
                               grepl("pfhpa", exposures) ~ "PFHpA",
                               grepl("pfhps", exposures) ~ "PFHpS",
                               grepl("pfhxs", exposures) ~ "PFHxS",
                               grepl("pfna", exposures) ~ "PFNA",
                               grepl("pfuna", exposures) ~ "PFUnA",
                               grepl("nmefosaa", exposures) ~ "NMeFOSAA"))%>%
  mutate(sig = ifelse(p.value < 0.05, "Sig", "Not Sig"))

cat_outcome_result1 <- model_output(pfas_names,
                                    outcomes = outcome[c(5,8,10)],
                                    covars = c("case1_control0", "sex", 
                                               "agebase", "tx",
                                               "est_creat_clear"),
                                    outcome_family = "binomial",
                                    data = data)

cat_result1 <- cat_outcome_result1 %>% 
  filter(term == "exposure_concentration") %>%
  select(-term)%>%
  filter(!exposures %in% 
           c("pfas_br_pfos", "pfas_l_pfoa", "pfas_l_pfos")) %>%
  mutate(pfas_name = case_when(grepl("pfos", exposures) ~ "PFOS",
                               grepl("pfoa", exposures) ~ "PFOA",
                               grepl("pfda", exposures) ~ "PFDA",
                               grepl("pfhpa", exposures) ~ "PFHpA",
                               grepl("pfhps", exposures) ~ "PFHpS",
                               grepl("pfhxs", exposures) ~ "PFHxS",
                               grepl("pfna", exposures) ~ "PFNA",
                               grepl("pfuna", exposures) ~ "PFUnA",
                               grepl("nmefosaa", exposures) ~ "NMeFOSAA")) %>%
  mutate(sig = ifelse(p.value < 0.05, "Sig", "Not Sig"))

(p3 <- cont_result1 %>%
  ggplot(aes(x = factor(pfas_name, levels = levels),y = estimate, color = sig)) +
  geom_point(size = 1) +
  geom_errorbar(aes(ymin = conf.low,
                    ymax = conf.high),
                width = 0) +
  geom_hline(yintercept = 0, linetype = 2) +
  facet_grid( ~ outcome, scales = "free") +
  ylab("Coefficient (95% CI)") +
  # ggtitle("Coefficient Plot (Binary Logistic Regression)") +
  theme(text = element_text(size = 10),
        axis.title.y = element_blank(),
        panel.background = element_rect(fill="white"), 
        strip.background = element_rect(fill = "white"),
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        legend.position = "none") +
  coord_flip() + 
  scale_color_manual(values = c("grey", "red")))

ggsave(fs::path(dir_figure,
                "1127_hw",
                "coef_plot_cont_outcome_full_group_with_extra_covar.png"), bg = "white",
       width =  8, height = 3, dpi = 300)

(p4 <- cat_result1 %>%
  ggplot(aes(x = factor(pfas_name, levels = levels),y = exp(estimate), color = sig)) +
  geom_point(size = 1) +
  coord_flip() +
  geom_errorbar(aes(ymin = exp(conf.low),
                    ymax = exp(conf.high)),
                width = 0) +
  geom_hline(yintercept = 1, linetype = 2) +
  facet_grid( ~ outcome, scales = "free") +
  ylab("Odds Ratio (95% CI)") +
  theme(text = element_text(size = 10),
        axis.title.y = element_blank(),
        panel.background = element_rect(fill="white"), 
        strip.background = element_rect(fill = "white"),
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        legend.position = "none") +
  coord_flip() + 
  scale_color_manual(values = c("grey", "red"))) 

ggsave(fs::path(dir_figure, 
                "1127_hw",
                "coef_plot_cat_outcome_full_group_with_extra_covar.png"), 
       bg = "white",
       width = 8, height = 3, dpi = 300)
```

## Time to event analysis
```{r survival analysis}
data <- full_data |> 
  tidylog::select(-all_of(prot_names), -contains("met_")) |>
  fastDummies::dummy_cols(select_columns = c('tx', 'sex'),
                          remove_selected_columns = TRUE, 
                          remove_most_frequent_dummy = TRUE) |>
  mutate(yrstohyp = (dxtime*12 + daystohyp)/365, 
         yrstohyp2 = yrstohyp + dxtime/12, 
         yrstomic = (daystomic)/365, 
         yrstomic2 = (yrstomic+dxtime/12),
         yrstomac = (daystomac)/365,
         yrstomac2 = (yrstomac+dxtime/12),
         # if_else(yrstomac > 8.5, 8.5, yrstomac), 
         mac2 = if_else(yrstomac > 8.5, 0, 1),
         mic2 = if_else(yrstomic > 8.5, 0, 1), 
         hyp2 = if_else(yrstohyp > 8.5, 0, 1), 
         across(all_of(pfas_names), scale))
```


