---
title: 'Today project'
author: "Hongxu Wang"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---
<style type="text/css">
body{
font-size: 12pt;
}
</style>

```{r exposure outcome setup, include=FALSE}
source(here::here("!libraries.R"))
source(here::here("!directories.R"))
source(here::here("!load_clean_data.R"))
source(here::here("!functions.R"))
source(here::here("analysis function_Multiple exposures and multiple outcomes analysis.R"))

options(knitr.table.format = "html")
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Examine key variables, set in the load_clean_data file.
```{r define variables}
# Glucose Outcomes: 
outcome_glu
# Tubular Injury Biomarkers 
outcome_biomaker 
# PFAS names for analysis
(analysis_pfas_names <- c(pfas_names_all, "score", "score_pfsas", "score_pfcas")) 
# Covariates for primary analysis
covars
# Order of PFAS for figures
levels
```

## Descriptive PFAS
```{r descriptive pfas}
# long format data for plasma pfas
data_l <- original_data %>% 
  pivot_longer(
    names_to = "pfas",
    values_to = "value",
    cols = c(all_of(pfas_names_all)))

# descriptive statistics
stats <- data_l %>%
  drop_na(value) %>%
  group_by(pfas) %>%
  dplyr::summarise(
    geometric_mean = fungm(value),
    percentile_50 = qntle_fxn(value, .50),
    percentile_75 = qntle_fxn(value, .75),
    percentile_90 = qntle_fxn(value, .9)) %>%
  ungroup()

# Rename columns
stats_summary <- stats %>%
  dplyr::rename(
    "PFAS" = pfas,
    "Geometric Mean [95% CI]" = geometric_mean,
    "50th percentile" = percentile_50,
    "75th percentile" = percentile_75,
    "90th percentile" = percentile_90)
# save result
# write_csv(stats_summary, fs::path(dir_results,
#                                   "1_0_descriptive_stats_pfas.csv"))
stats_summary
```

## Correlation of the PFAS
```{r correlation of pfas}
cor <- cor(data_scaled %>% dplyr::select(analysis_pfas_names))
corrplot::corrplot(cor)
```

## Exposure burden score
```{r exposure burden score}
# Exposure burden score now calculated in load clean data file
```


## Analysis with whole group 
**adjusted for "case1_control0", "sex", "agebase", "serum_creat"**
```{r whole group}

cont_outcome_result <- model_output(analysis_pfas_names,
                                    outcomes = c(outcome_glu, outcome_biomaker),
                                    covars = covars,
                                    outcome_family = "gaussian",
                                    data = data_scaled) %>%
  mutate(sig = ifelse(p.value < 0.05, "Sig.", "Not Sig."))

# cont_outcome_result <- epiomics::owas(df = data_scaled,
#                                       var = c(outcome_glu, outcome_biomaker),
#                                       omics = analysis_pfas_names,
#                                       covars = covars,
#                                       var_exposure_or_outcome = "outcome",
#                                       conf_int = TRUE)

cont_result <- cont_outcome_result %>% 
  filter(term == "exposure_concentration") %>%
  dplyr::select(-term) %>%
  rename(pfas = exposures) %>%
  rename_pfas() %>% 
  mutate(outcome_names = rename_outcomes(outcome))

## coefficient plot
(p1 <- cont_result %>%
    filter(outcome %in% c(outcome_glu)) %>%
    ggplot(aes(x = factor(pfas, levels = c(levels,"score")),
               y = estimate, 
               color = sig)) +
    geom_point(size = 1) +
    geom_errorbar(aes(ymin = conf.low,
                      ymax = conf.high),
                  width = 0) +
    geom_hline(yintercept = 0, linetype = 2) +
    facet_grid( ~ outcome_names, scales = "free") +
    ylab("Coefficient (95% CI)") +
    # ggtitle("Coefficient Plot (Binary Logistic Regression)") +
    theme(text = element_text(size = 10),
          axis.title.y = element_blank(),
          panel.background = element_rect(fill="white"), 
          strip.background = element_rect(fill = "white"),
          axis.line.x = element_line(color = "black"),
          axis.line.y = element_line(color = "black"),
          legend.position = "none") +
    coord_flip() +
    scale_color_manual(values = c("grey", "red")))

# ggsave(fs::path(dir_figure, 
#                 "1127_hw",
#                 "coef_plot_glucose_outcome.png"), bg = "white",
#        width = 8, height = 3, dpi = 300)

(p2 <- cont_result %>%
    filter(outcome %in% c(outcome_biomaker)) %>%
    ggplot(aes(x = factor(pfas, levels = c(levels,"score")),y = estimate, color = sig)) +
    geom_point(size = 1) +
    geom_errorbar(aes(ymin = conf.low,
                      ymax = conf.high),
                  width = 0) +
    geom_hline(yintercept = 0, linetype = 2) +
    facet_grid( ~ outcome_names, scales = "free") +
    ylab("Coefficient (95% CI)") +
    # ggtitle("Coefficient Plot (Binary Logistic Regression)") +
    theme(text = element_text(size = 10),
          axis.title.y = element_blank(),
          panel.background = element_rect(fill="white"), 
          strip.background = element_rect(fill = "white"),
          axis.line.x = element_line(color = "black"),
          axis.line.y = element_line(color = "black"),
          legend.position = "none") +
    coord_flip() +
    scale_color_manual(values = c("grey", "red")))

# ggsave(fs::path(dir_figure, 
#                 "1127_hw",
#                 "coef_plot_biomarker_outcome.png"), bg = "white",
#        width = 8, height = 3, dpi = 300)
```

## Survival Analysis -- Individual PFAS

```{r survival individual pfas}
## A. Set up for analysis ------
# Get the name of all PFAS exposure variables that were just created 
ind_vars <- data_scaled %>% 
  dplyr::select(
    contains("score"), 
    all_of(pfas_names_all)) |> 
  colnames() %>%  
  tibble(exposure = .)

# Get the name of all outcome variables created 
dep_vars <- tibble(time = c("daystohyp", "daystomic"),   
                   event = c("hyp", "mic")) 

# Get dataframe of all exposure outcome combinations
eo_comb <- list(pfas = ind_vars$exposure, event = dep_vars$event) %>%
  cross_df() %>% 
  left_join(dep_vars, by = "event", relationship = "many-to-many")

# Get the formula for all models
models <- eo_comb %>%
  mutate(covar = paste0("sex_male+",str_c(covars, collapse = "+")),
         formula = str_c("Surv(", time , ",", event, ")", "~", pfas, "+", covar))

## c. Run the models -----
models$output <- map(models$formula,
                     ~coxph(as.formula(.), data = data_scaled) %>%
                       tidy(., conf.int = TRUE))

# Clean up results
pfas_survival_models <- models %>%
  unnest(output) %>%
  tidylog::filter(grepl("score", term)| grepl("pfas_", term)| grepl("_sum", term)&!grepl("quintile",term)) %>%
  mutate(HR = exp(estimate),
         exp_conf_high = exp(conf.high),
         exp_conf_low = exp(conf.low),
         sig = ifelse(p.value < 0.05, "Sig.", "Not Sig.")) %>%
  rename_pfas() %>%
  mutate(pfas = ifelse(pfas %in% levels, pfas, term))
# Note: 1 unit increase in pfas resulting in xx-fold increase in hazard.


# Plot results
(p <- pfas_survival_models %>%
    ggplot(aes(x = term,y = estimate, color = sig)) +
    geom_point(size = 1) +
    geom_errorbar(aes(ymin = conf.low,
                      ymax = conf.high),
                  width = 0) +
    geom_hline(yintercept = 0, linetype = 2) +
    facet_grid( ~ time, scales = "free") +
    ylab("Log HR (95% CI)") +
    theme(text = element_text(size = 10),
          axis.title.y = element_blank(),
          panel.background = element_rect(fill="white"),
          strip.background = element_rect(fill = "white"),
          axis.line.x = element_line(color = "black"),
          axis.line.y = element_line(color = "black"),
          legend.position = "none") +
    coord_flip() +
    scale_color_manual(values = c("grey", "red")))

# ggsave(fs::path(dir_figure, 
#                 "1127_hw",
#                 "Hazard_ratio_burden_score_serum_creat.png"), bg = "white",
#        width = 8, height = 3, dpi = 300)
```

# JG Made it to here with updating the analysis based on the new covariates and burden scores

## Survival Analysis - Burden score(Binary) and survival outcome
```{r survival analysis}
# df = data_scaled %>% as.data.frame()
# df$score_median_named <- if_else(df$score_median == "0", 
#                                  "Low PFAS exposure", "High PFAS Exposure")
# 
# 
# # with case1_control0 covar
# fit <- coxph(Surv(yrstohyp2, hyp) ~ 
#                strata(score_median_named) + case1_control0 + sex + agebase + serum_creat, data = df)
# 
# # # without case1_control0 covar
# # fit <- coxph(Surv(yrstohyp2, hyp) ~ strata(score_median) + sex + agebase + serum_creat, data = df)
# # fit <- coxph(Surv(yrstoglyc2, glyc) ~ strata(score_median) + sex + agebase + serum_creat, data = df)
# # fit <- coxph(Surv(yrstomic2, mic) ~ strata(score_median) + sex + agebase + serum_creat, data = df)
# # summary(fit)
# 
# # Obtain dataset for the KM curves     
# curve <- surv_adjustedcurves(
#   fit = fit, 
#   variable = "score_median_named", 
#   data = df, 
#   reference = NULL, 
#   method = "conditional", 
#   size = 1) |>
#   rename(Years = time)
# 
# # Calculate incidence 
# curve$incidence <- (1-curve$surv)*100
# 
# # Make Plot 
# ggplot(curve, aes(x = Years, y = incidence, color = variable)) + 
#   geom_step(size = 1) + 
#   scale_y_continuous(limits =  c(0, 100)) + 
#   ylab("Hyperfiltration Incidence (%)") + 
#   scale_color_manual(values = c("black", "grey60"), 
#                      name = "PFAS Burden Score")  + 
#   scale_linetype_manual(values = c("solid","longdash"), 
#                         name = "PFAS Burden Score") + 
#   theme(legend.position = c(.1, .85)) 
# 
# # ggsave(filename = fs::path(dir_figure,"1127_hw", "KM_curve_score_hyp.jpeg"), 
# #        width = 5, 
# #        height = 5, dpi = 3000)
```


# Figure 1
```{r fig 1}
## 1A: Survival curves, dichotomous exposure burden score
(fig_1a <- ggplot(curve, aes(x = Years, 
                             y = surv, 
                             color = variable)) + 
   geom_step(size = 1) + 
   scale_y_continuous(limits =  c(0, 1)) + 
   ylab("Survival probability") + 
  scale_color_manual(values = c("black", "grey60"), 
                     name = "PFAS Burden Score")  + 
  scale_linetype_manual(values = c("solid","longdash"), 
                        name = "PFAS Burden Score") + 
  theme(legend.position = c(.1, .2), 
        text = element_text(size = 10)))


## fig 1B: individual PFAS -----
# filter to only key PFAS
all_pfas_hyp_res <-  pfas_survival_models |> 
  filter(str_detect(pfas, "score_", negate = TRUE), 
         time == "yrstohyp2") 

# reorder the pfas burden by the hazard ratio
all_pfas_hyp_res2 <-  all_pfas_hyp_res |> 
  mutate(pfas = if_else(str_detect(pfas, "score"), "PFAS Burden\nScore (Continuous)", pfas) %>% 
           fct_reorder(., estimate), 
         type = if_else(str_detect(pfas, "Burden"), 
                        "PFAS Burden Score (Continuous)", 
                        "Individual PFAS"), 
         sig2 = str_c(sig, type))


# Plot figure 1b- continuous PFAS effect estimates
(fig_1b <- all_pfas_hyp_res2 %>%
    ggplot(aes(x = pfas,y = estimate, color = sig)) +
    geom_point(size = 1) +
    geom_errorbar(aes(ymin = conf.low,
                      ymax = conf.high),
                  width = 0) +
    geom_hline(yintercept = 0, linetype = 2) +
    facet_grid(type ~ ., scales = "free_y", space = "free_y") +
    ylab("Log HR (95% CI)") +
    theme(text = element_text(size = 10),
          axis.title.y = element_blank(),
          panel.background = element_rect(fill="white"), 
          strip.text = element_blank(),
          axis.line.x = element_line(color = "black"),
          axis.line.y = element_line(color = "black"),
          legend.position = "none") +
    coord_flip() +
    scale_color_manual(values = c("grey70", "black")))


# Combine a and b
(figure_1 <- cowplot::plot_grid(NULL,  NULL, 
                   fig_1a, fig_1b, 
                   labels = c("A.", "B."),
                   rel_heights = c(.05, .9), 
                   rel_widths = c(.5, .5)))


ggsave(figure_1, filename = fs::path(dir_figure,"Figure 1 Hyperfiltration.jpeg"), 
       height = 5, width = 8, units = "in")
```


## Mixture analysis - qgcomp
```{r mixture analysis}
# qgcomp --------
qgcomp <- owas_qgcomp(df = data, 
                      expnms = analysis_pfas_names, 
                      omics = c(outcome_glu, outcome_biomaker), 
                      covars = covars,
                      confidence_level = 0.95,
                      q = 3) |>
  dplyr::select(feature, psi, lcl_psi, ucl_psi, p_value) %>%
  mutate(type = ifelse(feature %in% outcome_glu, "clinical", "biomarker")) %>%
  mutate(feature = rename_outcomes(feature))


reorder_outcome_name <- c("Hyperglycemia (HbA1c)", 
                          "β-cell function (C-Peptide oDI)",  
                          "Ins. Sensitivity (1/fast. ins)",
                          "NGAL","KIM-1", "IL18",
                          "FABPL","a-GST", "a-GST1","NAG"
) 
# Get color scale
colors <- RColorBrewer::brewer.pal(n=3, name = "Dark2")[1:2]

# Reorder outcome name and set color scale
qgcomp_res <- qgcomp %>%
  mutate(feature = factor(feature,
                          levels = reorder_outcome_name) %>%
           fct_rev(),
         color = if_else(type == "biomarker", 
                         colors[1], colors[2])) |>
  arrange(feature)

# Plot all outcomes 
(plotout <- qgcomp_res |>
    ggplot(aes(y = psi, 
               x = feature,
               color = color)) +
    geom_point(size = 3) +
    geom_errorbar(aes(ymin = lcl_psi,
                      ymax = ucl_psi),
                  width = 0, size = .75) +
    geom_hline(yintercept = 0, linetype = 2, linewidth = .5) +
    theme(axis.title.y = element_blank(), 
          axis.text.y = element_text(size = 14, vjust = .5)) +
    scale_color_manual(values = colors) +
    ylab("PFAS Mixture (95% CI)") + 
    coord_flip() + 
    theme(legend.position = "None",
          axis.text.y = element_text(color = qgcomp_res$color, 
                                     face = "bold")))

ggsave(plotout,
       filename = fs::path(dir_figure, 
                           "QGComp pfas mixture TODAY_1204.jpeg"),
       height = 2.25, width = 5, bg = "white")
```

## Survival analysis- Mixture PFAS
### glycemic failure
```{r glycemic failure}
(qc.survfit <- qgcomp::qgcomp.cox.noboot(
  survival::Surv(yrstoglyc2, glyc) ~ ., 
  expnms = analysis_pfas_names,
  data = data |> 
    dplyr::select(all_of(analysis_pfas_names), 
                  covars[-2],
                  sex_male,
                  yrstoglyc2, 
                  glyc),
  q = 2))
```
### hyperfiltration
```{r hyp}
(qc.survfit <- qgcomp::qgcomp.cox.noboot(
  survival::Surv(yrstohyp2, hyp) ~ ., 
  expnms = analysis_pfas_names,
  data = data |> 
    dplyr::select(all_of(analysis_pfas_names), 
                  covars[-2],
                  sex_male,
                  yrstohyp2, 
                  hyp),
  q = 2))
```

### microalbuminuria
```{r microalbuminuria}
(qc.survfit <- qgcomp::qgcomp.cox.noboot(
  survival::Surv(yrstomic2, mic) ~ ., 
  expnms = analysis_pfas_names,
  data = data |> 
    dplyr::select(all_of(analysis_pfas_names),
                  covars[-2],
                  sex_male,
                  yrstomic2, 
                  mic),
  q = 2))

```

## Survival Analysis - HbA1C and survival outcome

```{r survival analysis HbA1c}
df = data %>% as.data.frame()

#note: hba1c has already been normalized here.

fit <- coxph(Surv(yrstohyp2, hyp) ~ hb_a1c + sex + agebase + serum_creat, data = df)

# fit <- coxph(Surv(yrstoglyc2, glyc) ~ hb_a1c + case1_control0+ sex + agebase + serum_creat, data = df)

# fit <- coxph(Surv(yrstomic2, mic) ~ hb_a1c + case1_control0+ sex + agebase + serum_creat, data = df)

summary(fit)

# adjustedsurv(data = df,
#              variable = "score_median",
#              ev_time = "yrstohyp2",
#              event = "hyp",
#              method = "direct",
#              outcome_model = fit,
#              conf_int = TRUE, )
# ggadjustedcurves(fit,
#                  data = df,
#                  method = "conditional",
#                  variable = "score_median",
#                  # conf.int = TRUE,
#                  # pval = TRUE,
#                  # risk.table = TRUE,
#                  # legend.labs = c("Below Median", "Above Median"),
#                  legend.title = "PFAS Burden Score",
#                  palette = c("dodgerblue2", "orchid2"),
#                  title = "Kaplan-Meier Curve",
#                  # risk.table.height = 0.3
#                  )

# ggsave(filename = fs::path(dir_figure,"1127_hw", "KM_curve_score_hyp.jpeg"),
#        width = 5,
#        height = 5, dpi = 3000)
```

## Causal Mediation Analysis with Survival Outcome
* Exposure: PFAS burden score
* Mediator: HbA1c (normalized while in the analysis)
* Survival Outcome: Hyperfiltration (Kidney injury)
```{r CMA}
data1 <- data %>% mutate(score_median = factor(score_median))

set.seed(123)
mediation_result <- CMAverse::cmest(data1, 
                                    outcome = "yrstohyp2", 
                                    event = "hyp",
                                    exposure = "score", 
                                    mediator = c("hb_a1c"), 
                                    basec = NULL,
                                    EMint = TRUE, 
                                    mreg = list("linear"), 
                                    yreg = "coxph", 
                                    model = "gformula",
                                    astar = -1,
                                    a = 1,
                                    mval = list(0), 
                                    estimation = "imputation",
                                    inference = "bootstrap",
                                    nboot = 100
)

summary(mediation_result)

# CMAverse::ggcmest(mediation_result) +
#   ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 30, vjust = 0.8))
```

## Meet in the middle analysis
```{r meet in the middle analysis}
data_scale <- data %>% mutate_at(.vars = vars(omic_names),
                                 .funs = ~scale(.)%>%
                                   as.matrix())
## owas
result <- epiomics::owas(df = data_scale, 
                         var = "score",
                         omics = omic_names,
                         covars = covars,
                         var_exposure_or_outcome = "exposure",
                         family = "gaussian",
                         confidence_level = 0.95, 
                         conf_int = TRUE)

result_sig <- result %>% filter(p_value < 0.05)
dim(result_sig)

# try different kidney function: 

ind_vars <- tibble(exposure = colnames(data %>%
                                         dplyr::select(omic_names)))

dep_vars <- tibble(time = c("yrstohyp2"), 
                   event = c("hyp"))

eo_comb <- list(omic = ind_vars$exposure, event = dep_vars$event) %>%
  cross_df() %>% left_join(dep_vars, by = "event")

models <- eo_comb %>%
  mutate(covar = paste0("sex_male+",str_c(covars[-2], collapse = "+")),
         formula = str_c("Surv(", time , ",", event, ")", "~", omic, "+", covar))

models$output <- map(models$formula,
                     ~coxph(as.formula(.),
                            data = data_scale) %>%
                       tidy(., conf.int = TRUE))

mod_output_df <- models %>%
  unnest(output) %>%
  filter(grepl("seq", term)|grepl("met",term))%>%
  mutate(HR = exp(estimate),
         exp_conf_high = exp(conf.high),
         exp_conf_low = exp(conf.low),
         sig = ifelse(p.value < 0.05, "Sig.", "Not Sig."))

# Note: 1 unit increase in pfas resulting in xx-fold increase in hazard.
mod_output_df_sig <- mod_output_df %>% filter(sig == "Sig.")
dim(mod_output_df_sig)

## vocanol plot--
p <- mod_output_df %>%
  ggplot(aes(x = estimate,y = -log10(p.value), col = sig, label = term)) +
  geom_point(size = 1) +
  xlab("Log HR") +
  ylab("-log10(pvalue)") +
  theme(text = element_text(size = 10),
        panel.background = element_rect(fill="white"), 
        strip.background = element_rect(fill = "white"),
        axis.line.x = element_line(color = "black"),
        axis.line.y = element_line(color = "black"),
        legend.position = "none") +
  scale_color_manual(values = c("grey", "red"))

## meet in the middle
meet_in_the_middle_omics <- intersect(result_sig$feature_name, mod_output_df_sig$term)

meet_in_the_middle_omics

intersect(meet_in_the_middle_omics, outcome_biomaker)

sig_protein <- prot_metadata %>% filter(AptName %in% meet_in_the_middle_omics)

sig_protein

# write_csv(sig_protein, fs::path(dir_results, "protein_sig_meet_in_middle.csv"))
```

## High dimensional mediation analysis - HIMA
* Exposure: continous PFAS burden score or binary score
* Mediator: All metabolites and protein biomakers
* Survival Outcome: Hyperfiltration (Kidney injury)

```{r hima}
X1 = data$score %>% as.numeric()
# X1 = data$score_median %>% as.numeric()
Z1 = data[, c(covars[-2], "sex_male")] %>% as.matrix()

M = data %>% dplyr::select(contains("met"),contains("seq"),
                           -var_uni_value$name) %>%as.matrix()
M1 = data %>% dplyr::select(contains("met"), -var_uni_value$name) %>%as.matrix()
M2 = data %>% dplyr::select(contains("seq")) %>% as.matrix()

print("Both omics in the analysis")
(result_both<- HIMA::survHIMA(X = X1,
                              Z = Z1,
                              M = M,
                              OT = data$yrstohyp2,
                              status = data$hyp,
                              FDRcut = 1,
                              scale = TRUE,
                              verbose = FALSE))
print("Metabolites only")
(result_met<- HIMA::survHIMA(X = X1,
                             Z = Z1,
                             M = M1,
                             OT = data$yrstohyp2,
                             status = data$hyp,
                             FDRcut = 1,
                             scale = TRUE,
                             verbose = FALSE))
print("Protein only")
(result_pro<- HIMA::survHIMA(X = X1,
                             Z = Z1,
                             M = M2,
                             OT = data$yrstohyp2,
                             status = data$hyp,
                             FDRcut = 1,
                             scale = TRUE,
                             verbose = FALSE))


intersect(result_both$ID, outcome_biomaker)  
```
**Note: the FDRcut was set to as high as 1 here is to avoid error from the function. If FDRcut was set to 0.05 or 0.2. No significant result**
